### 2020-01-03 manuscript.Rmd
some updates

---
title: "Notebook with AMP-AD RNA-seq network analysis results"
output: html_notebook
---

Build aggregate modules from individual modules and differential expression meta-analysis, this will give you the same modules as are stored at syn11932957 - this code block builds the aggregate modules, and then checks if they are different from the paper aggregate modules.

```{r}
synapser::synLogin()
aggModsNew <- AMPAD::buildAggregateModules()
aggModsPaper <- synapser::synTableQuery('select * from syn11932957')$asDataFrame()[,-c(1,2)]

sum(aggModsNew[,1]!=aggModsPaper[,1])
sum(aggModsNew[,2]!=aggModsPaper[,2])
```

Make the pairwise enrichment plot (Figure 1B)

```{r}
synapser::synLogin()
AMPAD::pairwiseMatrixOfEnrichments('syn11932957')
```

Build aggregate modules without restricting the aggregation operation to the individual modules that are enriched for differentially expressed genes in AD.

```{r}
synapser::synLogin()
aggModsNonAdNew <- AMPAD::buildAggregateNonADModules()

```

Show overlap between AD and non AD (Figure SXF)
```{r}
AMPAD::compareADtoNonADmods(aggModsNew,
                            aggModsNonAdNew)

```

test building of gene sets

```{r}
targetedGeneSets <- AMPAD::collateEnrichmentSets()
```

Plot from Figure 1c
```{r}
synapser::synLogin()
modMeta <- AMPAD::getModuleMetainfo('syn11932957')
dummyDf <- targetedEnrichment$cell
dummyDf$fisherOR[dummyDf$adj.pval>=0.05] <- NA
#dummyDf <- dplyr::filter(targetedEnrichment$cell,adj.pval<=0.05)
dummyDf <- dplyr::left_join(dummyDf,modMeta)
dummyDf$category <- gsub('Zhang\\.','',dummyDf$category)
dummyDf$category <- factor(dummyDf$category,levels = rev(c('Astrocyte',
                                                       'Endothelial',
                                                       'Microglia',
                                                       'Neuron',
                                                       'MyelinOligos',
                                                       'NewOligos',
                                                       'OPC')))
g<-AMPAD::make_dot_plot(dummyDf,
                 xlab='AD Coexpression Module',
                 ylab='Cell Type Signature')
g
```